---
title: "Corona_Case_Prediction"
author: "Steven Smith, PhD"
date: "3/18/2020"
last_updated: "3/24/2020"
output: 
  html_document: 
    toc: yes
---

# The 2019-2020 Coronavirus Pandemic Analysis
## BACKGROUND & APPROACH
I wanted to track and trend the coronavirus outbreak on my own curiosity. There are some interesting questions that may fall out of this, as it is a very historic moment, including scientifically and analytically (we have a large amount of data being shared across the globe, analyzed in real-time). The world has come to a halt because of it.  
This analysis attempts to answer the following questions (more to come):
1. What does the trend of the pandemic look like to date?  
2. What are future case predictions based on historical model?
3. What interesting quirks or patterns emerge? 


ASSUMPTIONS & LIMITATIONS: 
* This data is limited by the source. I realized early on that depending on source there were conflicting # of cases. Originally I was using JHU data... but this was always 'ahead' of the Our World In Data. I noticed that JHU's website was buggy- you clicked on the U.S. stats but it didn't reflect the U.S.. So I changed data sources to be more consistent with what is presented in the media (and Our World In Data has more extensive plots I can compare my own to). An interesting aside might be why the discrepancy? Was I missing something?  
* Defintiions are important as is the idea that multiple varibales accumulate in things like total cases (more testing for example).  

SOURCE RAW DATA: https://ourworldindata.org/coronavirus
INPUT DATA LOCATION: github (https://github.com/sbs87coronavirus/data)
OUTPUT DATA LOCATIOn: github (https://github.com/sbs87coronavirus/results)

## PRE-ANALYSIS
The following sections are outside the scope of the 'analysis' but are still needed to prepare everything

### UPSTREAM PROCESSING/ANALYSIS - N/A
Process on a remote server
```{bash process_remote_server, eval=F}
# Typically code for processing data on a remote server. Used for tracability 

##------------------------------------------
# This denotes a subsection within the chunk
##------------------------------------------

##------------------------------------------

```

### SET UP ENVIORNMENT
Load libraries and set global variables
```{r setup, eval=T}

##------------------------------------------
## LIBRARIES
##------------------------------------------
library(ggplot2)
library(tidyverse)
library(plyr)
library(reshape2)
library(plot.utils)

##------------------------------------------

##------------------------------------------
# GLOBAL VARIABLES
##------------------------------------------
working_dir<-"/Users/ssmith/coronavirus/" # don't forget trailing /
results_dir<-paste0(working_dir,"results/") # assumes diretory exists
Corona_Cases.fn<-paste0(working_dir,"data/","total-cases-covid-19.csv") 
##------------------------------------------

```
### FUNCTIONS
List of functions
1. Function 1  
2. Function 2  
```{r functions, eval=F}
##------------------------------------------
## FUNCTION 1
##------------------------------------------
myfunction<-function(x){
  return(x+1)
}
##------------------------------------------
```

### READ IN DATA
* total number of cases, source: https://ourworldindata.org/coronavirus

```{r read_in_data, eval=T}
Corona_Cases<-read.csv(Corona_Cases.fn,header = T,stringsAsFactors = F)
```


### PROCESS DATA
* Rename column headers  
* Log10 transform total # cases  

```{r process, eval=T}

##------------------------------------------
## Rename column headers
##------------------------------------------
#Corona_Cases<-Corona_Cases %>% rename(replace = c("Total.confirmed.cases.of.COVID.19..cases."="Total_confirmed_cases","Year"="Days_Since_x"))
Corona_Cases<-Corona_Cases %>% rename(replace = c("Total.confirmed.cases.of.COVID.19..cases."="Total_confirmed_cases"))
library(lubridate)
#March 25 - they changed from # of days to date format
Corona_Cases$Date<-as.Date(Corona_Cases$Date,format = "%B %d, %Y")
Corona_Cases$Date.numeric<-as.numeric(Corona_Cases$Date)
as.Date(Corona_Cases$Date.numeric,origin = Corona_Cases$Date)
##------------------------------------------


##------------------------------------------
## log10 transform total # cases
##------------------------------------------
Corona_Cases$Total_confirmed_cases.log<-log(Corona_Cases$Total_confirmed_cases,10)
##------------------------------------------


```
## ANALYSIS
 

### Q1: What is the trend in total cases?
A blurb about the approach, some context
```{r question1, eval=T}

##------------------------------------------
## Compute # of days since 100th case
##------------------------------------------

plot_data<-filter(Corona_Cases,Code=="USA" & Total_confirmed_cases>0)
inflection<-min(plot_data[plot_data$Total_confirmed_cases>100,"Date.numeric"])
plot_data$Days_since_100<-plot_data$Date.numeric-inflection


#plot_data<-Corona_Cases[Corona_Cases$Code=="USA" & Corona_Cases$Days_Since_x>0,]
(model_fit<-lm(formula = Total_confirmed_cases.log~Days_since_100,data =plot_data[plot_data$Days_since_100>0,] ))
(slope<-model_fit$coefficients[2])
(intercept<-model_fit$coefficients[1])

# Formula for # of cases by x days
paste0("log10_total_cases = ",slope,"*days + ",intercept)
paste0("total_cases = 10^(",slope,"*days + ",intercept,")")
#Days untill... cases:
# 2.5k, 5k and 1M:
paste0("2.5k cases is ",(log(2.5E5,10) - intercept)/slope," days")
paste0("5k cases is ",(log(5E5,10)- intercept)/slope," days")
paste0("1M cases is ",(log(1E6,10)- intercept)/slope," days")


today<-max(plot_data$Days_since_100)
#tomorrow<-today+1
#tomorrow_plus1<-tomorrow+1
#tomorrow_plus2<-tomorrow+2
#tomorrow_plus3<-tomorrow+3
#next_week<-today+7
predicted_days<-today+c(1,2,3,7)

prediction_model<-function(m=1,b=0,days=1){
  total_cases.log<-m*days+b
  total_cases<-10^total_cases.log
  prediction<-data.frame(Days_since_100=days,Total_confirmed_cases=total_cases,Total_confirmed_cases.log=total_cases.log)
  return(prediction)
}
#today:
plot_data[plot_data$Days_since_100==today,]
prediction_model(m = slope,b=intercept,days=predicted_days)

cor(x = plot_data[plot_data$Days_since_100>0,"Days_since_100"],y = plot_data[plot_data$Days_since_100>0,"Total_confirmed_cases.log"])


(total_cases.25.plot<-ggplot(plot_data,aes(x=Days_since_100,y=Total_confirmed_cases))+
  default_theme+
  geom_point())

(total_cases.25.log.plot<-ggplot(plot_data,aes(x=Days_since_100,y=Total_confirmed_cases.log))+
  geom_abline(slope = slope,intercept = intercept,lty=2)+
   default_theme+
  geom_point())

write_plot(total_cases.25.plot,wd = results_dir)
write_plot(total_cases.25.log.plot,wd = results_dir)



```
### What is the predicted number of cases?


# What is the prediction of COVID-19 based on model thus far?
Additional questions:

WHy did it take to day 40 to start a log linear trend?
How long will it be till x number of cases?
When will the plateu happen?
Are any effects noticed with social distancing? Delays
```{r prediction, eval=F}

plot_data$type<-"Historical"
names(Corona_Cases)

#Corona_Cases_wprediction<-rbind.fill(Corona_Cases,data.frame(Code="USA",type="MAR25_prediction",prediction_model(m=slope,b=intercept,days = predicted_days)))
tmp<-Corona_Cases_wprediction
Corona_Cases_wprediction<-rbind.fill(plot_data,data.frame(Code="USA",type="MAR26_prediction",prediction_model(m=slope,b=intercept,days = predicted_days)))

plot_data.prediction<-Corona_Cases_wprediction #filter(Corona_Cases_wprediction,Code=="USA" )
prediction_values<-prediction_model(m=slope,b=intercept,days = predicted_days)$Total_confirmed_cases

mar26_prediction<-data.frame(type="MAR26_prediction",Days_since_100=predicted_days-7,Total_confirmed_cases.log=log(prediction_values,10),text_display=paste0(c("tomorrow : ","two_days: ","three_days: ","next_week: "),round(prediction_values,digits = 0)))
mar25_prediction<-data.frame(type="MAR25_prediction",Days_since_100=predicted_days-20,Total_confirmed_cases.log=log(c(56490.42,75400.32,100640.22,319422.18),10),text_display=paste0(c("today's (actual=55231): ","tomorrow: ","two_days: ","6 days: "),round(c(56490.42,75400.32,100640.22,319422.18),digits = 0)))
              
(predicted.25.plot<-ggplot(plot_data.prediction,aes(x=Days_since_100,y=Total_confirmed_cases.log,col=type))+
    geom_abline(slope = slope,intercept = intercept,lty=2)+
  #geom_vline(xintercept = c(key_milestones.days),lty=2)+
  geom_point(size=3)+
default_theme+
  ylab("log10 Total U.S. Cases")+
  xlab("Days since X")+
  theme(legend.position = "bottom")+
    geom_text(data = mar25_prediction,aes(label=text_display))+
        geom_text(data = mar26_prediction,aes(label=text_display))+
  scale_color_manual(values = c("Historical"="#377eb8","MAR25_prediction"="orange","MAR26_prediction"="#e41a1c")))
  
write_plot(predicted.25.plot,wd = results_dir)

```


```{r question2, eval=F}

##------------------------------------------
## filter input_data1
##------------------------------------------
input_data1.filter<-fitler(input_data1,col1=="foo")
##------------------------------------------

##------------------------------------------
## sub question 1
##------------------------------------------
table(input_data1.filter$col<5)
##------------------------------------------

##------------------------------------------
## sub question 2
##------------------------------------------
table(input_data1.filter$col<10)
##------------------------------------------

##------------------------------------------
## plot data
##------------------------------------------
(input_data1.filter.plot<-ggplot(input_data1.filter,aes(x=col1,y=col2.log))+
   geom_point()+
   default_plot_theme)
write_plot(input_data1.filter.plot,wd=results_dir)
##------------------------------------------

```
# CONCLUSION
A concluding remark(s) on the major findings, preferabbly to pointers where the data can be found. 

Helps to have a bullet point for each analysis chunk or an answer to each of the above 'questions':
*  Answer 1. 
*  Answer 2.  

#END
Cheatsheet:
http://rmarkdown.rstudio.com>
# TODO
* mkdir the results dir if it doesn't exist
* make ggplot a dependency for plot.utils?
